### 2019-03-22 Journal Version/gap.tex
journal

\section{Stronger version of Carr-Vempala for covering problems} \label{gap}

@@

	Let $u^i(\epsilon) = \frac{Cv^i-\epsilon z^i}{C(1-\epsilon)}$. For $j\in \{1,\ldots,m\}$, let $t_j = A_jz^i-Cb$ and $s_j=A_jv^i-b$. 
	Assume $\epsilon <1$. For $j\in \{1,\ldots,m\}$ such that $A_jz^i\leq Cb$, i.e. $t_j\leq 0$ (this includes $j$ such that $A_jv^i=b$, i.e. $s_j=0$) we have
	\begin{equation}
	A_j u^i(\epsilon) = \frac{CA_jv^i-\epsilon A_jz^i}{C(1-\epsilon)}\geq \frac{Cb-\epsilon Cb}{C(1-\epsilon)}=b
	\end{equation}
	Assume $\epsilon \leq C\cdot\min_{j: t_j >0} \frac{s_j}{t_j}$. Notice that $\max_{j: t_j >0} \frac{s_j}{t_j}>0$ since for $j$ such that $t_j>0$, we have $A_jz_i > Cb$ which implies that $A_jv^i>b$, so $s_j>0$. For $j$ such that $t_j>0$ we have
	\begin{align*}
	A_ju^i(\epsilon) &= \frac{CA_jv^i-\epsilon A_jz^i}{C(1-\epsilon)}& \\
	&= \frac{C(b+s_j)-\epsilon (Cb+t_j)}{C(1-\epsilon)}&\\
	&= b+ \frac{Cs_j-\epsilon t_j}{C(1-\epsilon)}&\\
	&\geq b & (\epsilon  \leq C\cdot  \frac{s_j}{t_j}, \mbox{ and } \epsilon <1, \mbox{ so } \frac{Cs_j-\epsilon t_j}{C(1-\epsilon)}\geq 0 )
	\end{align*} 
	Therefore if $\epsilon \leq 1$ and $\epsilon \leq C\cdot\min_{j: t_j >0} \frac{s_j}{t_j}$ we have $Au^i(\epsilon)\geq \textbf{1}b$. Next, we show that we can choose $\epsilon$ so that $ u^i(\epsilon)\geq 0$.

@@

	We proceed by induction on $i$. If $i=1$, we just need to show that $Cv^1$ is dominated by a convex combination of $z^1$ and $Cv^2,\ldots,Cv^r$. By Lemma \ref{epsilon}, there exists $x^1\leq  \frac{Cv^1-\epsilon_1z^1}{C(1-\epsilon_1)}$. We have $x^1\in P$, so we can write $x^1$ as convex combination of $v^1,\ldots, v^r$. We have
	\begin{equation}
	Cv^1 \geq  \epsilon_1z^1 + C(1-\epsilon_1)x^1 \geq \epsilon_1z^1 + C(1-\epsilon_1) \sum_{\ell=1}^{r}\lambda^1_\ell v^\ell.
	\end{equation}
	Hence,
	\begin{equation}
	Cv^1 \geq \frac{\epsilon_1}{1-(1-\epsilon_1)\lambda^1_1}\cdot z^1 + (1-\epsilon_i)\sum_{\ell=2}^{r} \frac{\lambda^1_\ell}{1-(1-\epsilon_1)\lambda_1^1}\cdot Cv^\ell.
	\end{equation}
	Observe that $\frac{\epsilon_1}{1-(1-\epsilon_1)\lambda^1_1}$ and $(1-\epsilon_1)\frac{\lambda^1_\ell}{1-(1-\epsilon_1)\lambda_1^1}$ for $\ell=2,\ldots,r$ form convex multipliers, so the base case holds.

@@

>>>>

\section{Strengthening Theorem \ref{CV} for bounded covering problems} \label{gap}

@@

	Let $u^i(\epsilon) = \frac{Cv^i-\epsilon z^i}{C(1-\epsilon)}$. For $j\in \{1,\ldots,m\}$, let $t_j = A_jz^i-Cb$ and $s_j=A_jv^i-b$. 
	Assume $\epsilon <1$. For $j\in \{1,\ldots,m\}$ such that $A_jz^i\leq Cb$, i.e. $t_j\leq 0$ (this includes $j$ such that $A_jv^i=b$, i.e. $s_j=0$) we have
	\begin{equation*}
	A_j u^i(\epsilon) = \frac{CA_jv^i-\epsilon A_jz^i}{C(1-\epsilon)}\geq \frac{Cb-\epsilon Cb}{C(1-\epsilon)}=b
	\end{equation*}
	Assume $\epsilon \leq C\cdot\min_{j: t_j >0} \frac{s_j}{t_j}$. Notice that $\max_{j: t_j >0} \frac{s_j}{t_j}>0$ since for $j$ such that $t_j>0$, we have $A_jz_i > Cb$ which implies that $A_jv^i>b$, so $s_j>0$. For $j$ such that $t_j>0$ we have
	\begin{align*}
	A_ju^i(\epsilon) &= \frac{CA_jv^i-\epsilon A_jz^i}{C(1-\epsilon)}& \\
	&= \frac{C(b+s_j)-\epsilon (Cb+t_j)}{C(1-\epsilon)}&\\
	&= b+ \frac{Cs_j-\epsilon t_j}{C(1-\epsilon)}&\\
	&\geq b & (\epsilon  \leq C\cdot  \frac{s_j}{t_j}, \mbox{ and } \epsilon <1, \mbox{ so } \frac{Cs_j-\epsilon t_j}{C(1-\epsilon)}\geq 0 )
	\end{align*} 
	Therefore if $\epsilon \leq 1$ and $\epsilon \leq C\cdot\min_{j: t_j >0} \frac{s_j}{t_j}$ we have $Au^i(\epsilon)\geq \textbf{1}b$. Next, we show that we can choose $\epsilon$ so that $ u^i(\epsilon)\geq 0$.

@@

	We proceed by induction on $i$. If $i=1$, we just need to show that $Cv^1$ is dominated by a convex combination of $z^1$ and $Cv^2,\ldots,Cv^r$. By Lemma \ref{epsilon}, there exists $x^1\leq  \frac{Cv^1-\epsilon_1z^1}{C(1-\epsilon_1)}$. We have $x^1\in P$, so we can write $x^1$ as convex combination of $v^1,\ldots, v^r$. We have
	\begin{equation*}
	Cv^1 \geq  \epsilon_1z^1 + C(1-\epsilon_1)x^1 \geq \epsilon_1z^1 + C(1-\epsilon_1) \sum_{\ell=1}^{r}\lambda^1_\ell v^\ell.
	\end{equation*}
	Hence,
	\begin{equation}\label{equation6}
	Cv^1 \geq \frac{\epsilon_1}{1-(1-\epsilon_1)\lambda^1_1}\cdot z^1 + (1-\epsilon_i)\sum_{\ell=2}^{r} \frac{\lambda^1_\ell}{1-(1-\epsilon_1)\lambda_1^1}\cdot Cv^\ell.
	\end{equation}
	Observe that $\frac{\epsilon_1}{1-(1-\epsilon_1)\lambda^1_1}$ and $(1-\epsilon_1)\frac{\lambda^1_\ell}{1-(1-\epsilon_1)\lambda_1^1}$ for $\ell=2,\ldots,r$ form convex multipliers, so the base case holds.

@@

\begin{example}

Consider  $\EC(G) = \{x\in [0,2]^E_{\geq 0}: \; x(\delta(U))\geq 2 \mbox{ for $\emptyset \subset U \subset V$}\},$ where $G=(V,E)$ is the graph in Figure \ref{example}. 
\begin{figure}[h]
	\centering
	\includegraphics[width=0.2\linewidth]{example.pdf}
	\caption{Graph $G=(V,E)$. Let $H$ be the Hamiltonian cycle of $G$ that contains edges $(1,2),(2,3),(3,4),(4,5),(5,6),(6,7),(7,8),(1,8)$. Let $M= E\setminus H$: $M=\{(1,5),(2,7),(3,6),(4,8)\}$.}
	\label{example}
\end{figure}

Let $v^1$ be the following extreme point of $\EC(G)$: \begin{equation*}
v^1_{e} = \left\{ \begin{array}{ll}
1 & \mbox{if $e\in M$};\\
1/2 & \mbox{if $e\in H$}.
\end{array} \right. 
\end{equation*}
Define $z^1$ to be the following vector:
\begin{equation*}
z^1_{e} = \left\{ \begin{array}{ll}
7/5 & \mbox{if $e= (1,5)$};\\
1 & \mbox{if $e\in M\setminus \{(1,5)\}$};\\
2/5 & \mbox{if $e \in \{(1,8),((4,5))\}$};\\
3/5 & \mbox{if $e\in H\setminus\{(1,8),((4,5))\} $}.
\end{array} \right. 
\end{equation*}
It is easy to check that $z^1$ satisfies the requirement of Theorem \ref{gap} when setting $C=\frac{6}{5}$. In particular, we have $ z^1(\delta(U))\leq \frac{6}{5}v^1(\delta(U))$ for all $\emptyset\subset U \subset V$. Also, is it easy to decompose $z^1$ into a convex combination of integer point in of $\EC(G)$.

Now, we want to set the $\epsilon_i$ in the statement of Lemma \ref{epsilon} to be $2^{-k}$. By Lemma \ref{epsilon}
\begin{equation*}
x^1_{e} = \left\{ \begin{array}{ll}
\frac{2^k-7/6}{2^k-1} & \mbox{if $e= (1,5)$};\\
\frac{2^k-5/6}{2^k-1} & \mbox{if $e\in M\setminus \{(1,5)\}$};\\
\frac{2^k- 2/3}{2\cdot (2^k-1)} & \mbox{if $e \in \{(1,8),((4,5))\}$};\\
\frac{1}{2} & \mbox{if $e\in H\setminus\{(1,8),((4,5))\} $}.
\end{array} \right. 
\end{equation*}
is in $\EC(G)$ for sufficiently large $k$. The vectors defined below $v^2,\ldots,v^7$ are also extreme points of $\EC(G)$.

\begin{table}[h]
	\centering
	\begin{tabular}{ccccccccccccc}
		\toprule
		& (1,5) & (2,7)& (3,6)&(4,8)&(1,2)&(2,3)&(3,4)&(4,5)&(5,6)&(6,7)&(7,8)&(1,8)      \\ \midrule
			$v^2=$ & (0&1&2&1&0&1&1&2&0&0&1&2)  \\
			$v^3=$ & (0&2&2&1&0&0&0&1&1&1&1&2)\\
			$v^4=$ & (1&2&2&2&1&1&1&1&0&0&0&0)\\
			$v^5=$ & (0&2&1&1&1&1&0&1&1&0&0&1)\\
			$v^6=$ & (1&1&1&2&0&1&0&0&1&0&1&1)\\
			$v^7=$ & (1&1&1&2&1&0&1&1&0&1&0&0)\\
				 \bottomrule
	\end{tabular}
\end{table}
Moreover, we have 
\begin{equation*}
x^1 = (1-8\lambda)v^1+ \lambda v^2 + 2\lambda v^3+\lambda v^4+ \lambda v^5+ \lambda v^6 + 2\lambda v^7, \quad \lambda = \frac{1}{24\cdot (2^k-1)}
\end{equation*}
Now, this allows us to rewrite Inequality \ref{equation6} for this example.
\begin{equation*} 
\frac{6}{5}v^1 \geq \frac{3}{4} z^1 +\frac{1}{32}(\frac{6}{5}\cdot v^2) +\frac{1}{16}(\frac{6}{5}\cdot v^3) + \frac{1}{32}(\frac{6}{5}\cdot v^4)+\frac{1}{32}(\frac{6}{5}\cdot v^5) + \frac{1}{32}(\frac{6}{5}\cdot v^6)+ \frac{1}{16}(\frac{6}{5}\cdot v^7).  
\end{equation*}
\end{example}

### 2019-03-22 Journal Version/introduction.tex
journal

Mixed-integer linear programming (MILP), the optimization of a linear objective function subject to linear and integrality constraints, models many practical optimization problems including scheduling, logistics and resource allocation.  The set of feasible points for a MILP is the set
\begin{equation}
S(A,G,b)= \{(x,y)\in \bbbz^{n}\times \bbbr^p\;:\; Ax+Gy\geq b\}  \label{S}.
\end{equation}
If we drop the integrality constraints, we have the linear relaxation of set $S(A,G,b)$,
\begin{equation}
P(A,G,b) = \{(x,y)\in \bbbr^{n+ p}\;:\; Ax+Gy\geq b\}. \label{P}
\end{equation}

Let $I=(A,G,b)$ be the feasible set of a specific instance. Then $S(I)$ and $P(I)$ denote $S(A,G,b)$ and $P(A,G,b)$, respectively. Given a linear objective function $c$, a MILP problem is $\min \;\{cx:\; (x,y) \in S(I)\}$. It is  NP-hard even to determine if a MILP instance has a feasible solution~\cite{GJ79}. However, intelligent branch-and-bound strategies allow commercial and open-source MILP solvers to give exact solutions (or near-optimal with provable bound) to many specific instances of NP-hard combinatorial optimization problems.

@@

Many researchers (see \cite{vazirani,sw}) have developed polynomial-time LP-based algorithms that find solutions for special classes of MILPs whose cost are provably smaller than $C\cdot z_{LP}(I,c)$. The approximation factor $C$ can be a constant or depend a MILP parameter, e.g. $O(\log(n))$. However, for many combinatorial optimization problems there is a limit to such techniques. Define the \textit{integrality gap} of the MILP formulation for instance $I$ to be $g_I= \max_{c\geq 0}\frac{z_{IP}(I,c)}{z_{LP}(I,c)}$. This value depends on the constraints in (\ref{S}).  We cannot hope to find solutions for the MILP with objective values better than $g_I\cdot z_{LP}(I,c)$.

@@

We give a general approximation framework for solving $\{0,1\}$-MILPs.  Consider the set of point described by sets $S(I)$ and $P(I)$ as in (\ref{S}) and (\ref{P}), respectively. Assume in addition that $S(I),P(I)\subseteq [0,1]^n\times \bbbr^p$. For a vector $x\in \bbbr^n$ such that $(x,y)\in P(I)$ for some $y\in \bbbr^p$, let $\spp(x)= \{i \in \{1,\ldots,n\}: x_i \neq 0\}$.

\textit{Fractional Decomposition Tree} (FDT) is a polynomial-time algorithm that given a point $x\in P(I)$ produces a convex combination of feasible points in $S(I)$ that are dominated by a ``factor" $C$ of $x$. If $C = g_I$, it would be optimal. However we can only guarantee a factor of $g_I^{|\spp(x)|}$. FDT relies on iteratively solving linear programs that are about the same size as the description of $P(I)$.

\begin{restatable}{theorem}{binaryFDT}
	\label{binaryFDT}
	Assume $1\leq g_I 	<\infty$. 	
	The Fractional Decomposition Tree (FDT) algorithm , given $(x^*,y^*)\in P(I)$, produces in polynomial time $\lambda\in [0,1]^k$ and $(z^1,w^1),\ldots,(z^k,w^k) \in S(I)$ such that $k\leq |\spp(x^*)|$, $\sum_{i=1}^{k}\lambda_i z^i\leq Cx^*$, and $\sum_{i=1}^{k}\lambda_i = 1$. Moreover, $C\leq g_I^{|\spp(x^*)|}$.
\end{restatable}

@@

We consider the following TSP-related problems.  The {\em 2-edge-connected subgraph problem (2EC)} is to find a minimum-weight 2-edge-connected (multigraph) subgraph
% (which can multiple copies of each edge)
in a graph $G=(V,E)$ with respect to weights $c\in \bbbr^E_{\geq 0}$. In the {\em tree-augmentation problem (TAP)} we wish to add a minimum-cost set of edges to a tree to make it 2-edge-connected.  We formally define TAP in Section~\ref{experiment}.

One can extend the FDT algorithm for binary MILPs into covering $\{0,1,2\}$-MILPs by losing a factor $2^{|\spp(x)|}$. In order to eradicate this factor, we need to treat the coordinate $i$ with $x_i=1$ differently. The 2EC problem has the natural linear programming relaxation is
%\begin{equation}\label{2ECpol}
$\EC(G)= \{x\in \bbbr_{\geq 0}^{E}\;:\; x(\delta(S))\geq 2 \text{ for $\emptyset \subset S\subset V$})\}.$
%\end{equation}
% We prove the following theorem.

\begin{restatable}{theorem}{FDTEC}
	\label{FDT2EC}
	Let $G=(V,E)$ and $x$ be an extreme point of  $\EC(G)$. The FDT algorithm for 2EC produces $\lambda\in [0,1]^k$ and 2-edge-connected multigraphs $F_1,\ldots,F_k$ such that $k\leq 2|V|-1$, $\sum_{i=1}^{k}\lambda_i \chi^{F_i}\leq Cx^*$, and $\sum_{i=1}^{k}\lambda_i = 1$. Moreover, $C\leq g({\EC})^{k}$, where $g({\EC})$ is the integrality gap of the 2-edge-connected multigraph problem with respect to formulation in (\ref{2ECpol}) 
\end{restatable}

@@

Known polyhedral structure makes it easier to study integrality gaps for such problems. Carr and Ravi \cite{carr-ravi} introduced fundamental extreme points. A point x in Held-Karp relaxation for TSP (or 2EC; they have the same relaxation) is a point whose support of $x$, namely $G_x$ satisfies the following: i)  $G_x$ is a cubic graph, ii) in $G_x$ there is exactly one edge with $x_e=1$ incident to each node iii) The fractional edges of $G_x$ form a Hamiltonian cycle.  We say a fundamental extreme point (FEP) is {\em order $k$} if there are $k$ nodes on this Hamiltonian cycle. An FEP of order $k$ could represent an instance with many more than $k$ vertices. Carr and co-authors~\cite{CV,carr-ravi,BC11} proved that showing that $Cx$ is a convex combination of tours (resp. 2-edge-connected subgraphs) for all fundamental extreme points is equivalent to proving that the integrality gap for TSP (resp. 2EC) is bounded above by $C$. Thus we can create the ``hardest'' LP solutions to decompose.

@@

\paragraph{Contributions}  The paper has the following contributions: 
\begin{itemize}
\item We give a simple algorithm, Dom2IP, that can prove a formulation’s integrality gap is unbounded or if not provide a feasible integer solution.  Someone formulating a first MILP for a new problem can test it with Dom2IP.  If the algorithm ever fails in finding a feasible solution, the MILP has an unbounded gap.
\item We give an algorithm, Fractional Decomposition Tree (FDT), to construct the convex decomposition in the Carr-Vempala theorem, perhaps scaling by a factor larger than the integrality gap.  Each step of this algorithm provides a {\em lower bound} on the instances integrality gap. This also provides a lower bound on the approximation factor of any LP-based approximation algorithm using this formulation. The overall approximation factor of the FDT algorithm is an upper bound on the integrality gap for that specific instance.
\item For a special set of problems related to TSP, where there is a notion of a fundamental extreme point and long-running attempts to exactly determine the integrality gap of classic formulations, experimental analysis with FDT can help give some intuition about which bound(s) is/are likely to be loose.  Computing on fundamental extreme points is a way to experimentally characterize the gap upper bound.  There is no guarantee. Still, this can help direct theoretical analysis in the most promising direction.  Furthermore, for these kinds of problems, especially 2EC, FDT gives good approximate solutions, better than the best current competitor (Christofides).
\end{itemize}

\iffalse{
Finally we give a stronger characterization of integrality gap than that in Theorem \ref{CV} for bounded covering problems. For this purpose assume $P= \{x\in \bbbr^n_{\geq 0}: Ax\geq b\textbf{1}, x \leq b\textbf{1}\}$, $S= P\cap \bbbz^n$, and $g= \max_{c\geq 0} \frac{\min_{x\in S}cx}{\min_{x\in P}cx}$. Examples of such problems include the 2-edge-connected multigraph problem, the tree augmentation problem, and many others.

@@

This means in order to prove an upper bound on the integrality gap of a covering problem, we need to show there is a convex combination of integer feasible points that is ``cheap'' on all tight cuts. Notice that Theorem \ref{CV} requires the certificate convex combination to be ``cheap" on every single variable.
\subsection{Notations}
For vectors $x,y\in \bbbr_{n}$ we say $x$ dominates $y$ if $x_i\geq y_i$ for $i= 1,\ldots,n$. For $m\times n$ matrix $A$, let $A_j$ be the $j$-th row of $A$ and $A^j$ be the $j$-th column of $A$. Let $\textbf{1}$ be the vector of all ones of a suitable dimension. For a set $S$ of vectors in $\bbbr_{n}$, $\conv(S)$ is the convex hull of all the points in $S$.
}\fi

>>>>

Mixed-integer linear programming (MILP), the optimization of a linear objective function subject to linear and integrality constraints, models many practical optimization problems including scheduling, logistics and resource allocation.  The set of feasible points for a MILP is the set
\begin{equation}
S(A,G,b)= \{(x,y)\in \mathbb{Z}^{n}\times \mathbb{R}^p\;:\; Ax+Gy\geq b\}  \label{S}.
\end{equation}
If we drop the integrality constraints, we have the linear relaxation of set $S(A,G,b)$,
\begin{equation}
P(A,G,b) = \{(x,y)\in \mathbb{R}^{n+ p}\;:\; Ax+Gy\geq b\}. \label{P}
\end{equation}

Let $I=(A,G,b)$ be the feasible set of a specific instance. Then $S(I)$ and $P(I)$ denote $S(A,G,b)$ and $P(A,G,b)$, respectively. Given a linear objective function $c$, a Mixed-Integer-Linear Program (MILP) is $\min \;\{cx:\; (x,y) \in S(I)\}$. It is  NP-hard even to determine if an MILP instance has a feasible solution~\cite{GJ79}. However, intelligent branch-and-bound strategies allow commercial and open-source MILP solvers to give exact solutions (or near-optimal with provable bound) to many specific instances of NP-hard combinatorial optimization problems.

@@

Many researchers (see \cite{vazirani,sw}) have developed polynomial-time LP-based algorithms that find solutions for special classes of MILPs whose cost are provably smaller than $C\cdot z_{LP}(I,c)$. The approximation factor $C$ can be a constant or depend a parameter of the MILP, e.g. $O(\log(n))$. However, for many combinatorial optimization problems there is a limit to such techniques. Define the \textit{integrality gap} of the MILP formulation for instance $I$ to be $g_I= \max_{c\geq 0}\frac{z_{IP}(I,c)}{z_{LP}(I,c)}$. This value depends on the constraints in (\ref{S}).  We cannot hope to find solutions for the MILP with objective values better than $g_I\cdot z_{LP}(I,c)$.

@@

We give a general approximation framework for solving $\{0,1\}$-MILPs.  Consider the set of point described by sets $S(I)$ and $P(I)$ as in (\ref{S}) and (\ref{P}), respectively. Assume in addition that $S(I),P(I)\subseteq [0,1]^n\times \mathbb{R}^p$. For a vector $x\in \mathbb{R}_{\geq 0}^n$ such that $(x,y)\in P(I)$ for some $y\in \mathbb{R}^p$, let $\spp(x)= \{i \in \{1,\ldots,n\}: x_i \neq 0\}$.

\textit{Fractional Decomposition Tree} (FDT) is a polynomial-time algorithm that given a point $(x,y)\in P(I)$ produces a convex combination of feasible points in $S(I)$ that are dominated by a ``factor" $C$ of $x$ in the coordinates corresponding to $x$. If $C = g_I$, it would be optimal. However we can only guarantee a factor of $g_I^{|\spp(x)|}$. FDT relies on iteratively solving linear programs that are about the same size as the description of $P(I)$.

\begin{restatable}{theorem}{binaryFDT}
	\label{binaryFDT}
	Assume $1\leq g_I 	<\infty$. 	
	The Fractional Decomposition Tree (FDT) algorithm, given $(x^*,y^*)\in P(I)$, produces in polynomial time $\lambda\in [0,1]^k$ and $(z^1,w^1),\ldots,(z^k,w^k) \in S(I)$ such that $k\leq |\spp(x^*)|$, $\sum_{i=1}^{k}\lambda_i z^i\leq Cx^*$, and $\sum_{i=1}^{k}\lambda_i = 1$. Moreover, $C\leq g_I^{|\spp(x^*)|}$.
\end{restatable}

@@

We consider the following TSP-related problems.  The {\em 2-edge-connected subgraph problem (2EC)} is to find a minimum-weight 2-edge-connected multigraph (subgraph which can contain multiple copies of each edge)
in a graph $G=(V,E)$ with respect to weights $c\in \mathbb{R}^E_{\geq 0}$. In the {\em tree-augmentation problem (TAP)} we wish to add a minimum-cost set of edges to a tree to make it 2-edge-connected.  We formally define TAP in Section~\ref{experiment}.

One can extend the FDT algorithm for binary MILPs into covering $\{0,1,2\}$-MILPs by losing a factor $2^{|\spp(x)|}$ on top of the loss for FDT. In order to eradicate this extra factor, we need to treat the coordinate $i$ with $x_i=1$ differently. For 2EC we are able to achieve this. The 2EC problem has the natural linear programming relaxation is
\begin{equation}\label{2ECpol}
\EC(G)= \{x\in [0,2]^{E}\;:\; x(\delta(U))\geq 2 \text{ for $\emptyset \subset U\subset V$})\}.
\end{equation}
 We prove the following theorem.

\begin{restatable}{theorem}{FDTEC}
	\label{FDT2EC}
	Let $G=(V,E)$ and $x$ be an extreme point of  $\EC(G)$. The FDT algorithm for 2EC produces $\lambda\in [0,1]^k$ and 2-edge-connected multigraphs $F_1,\ldots,F_k$ such that $k\leq 2|V|-1$, $\sum_{i=1}^{k}\lambda_i \chi^{F_i}\leq Cx^*$, and $\sum_{i=1}^{k}\lambda_i = 1$. Moreover, $C\leq g_{\EC}^{k}$, where $g_{\EC}$ is the integrality gap of the 2-edge-connected multigraph problem with respect to formulation in (\ref{2ECpol}) 
\end{restatable}

We give a stronger characterization of integrality gap than that in Theorem \ref{CV} for bounded covering problems. For this purpose assume $P= \{x\in \mathbb{R}^n_{\geq 0}: Ax\geq b\cdot \textbf{1}, x \leq b\cdot \textbf{1}\}$, where $A\in \mathbb{Z}^{m\times n}_{\geq 0}$ and $b\in \mathbb{Z}_{\geq 0}$. Let $S= P\cap \mathbb{Z}^n$ and $g= \max_{c\geq 0} \frac{\min_{x\in S}cx}{\min_{x\in P}cx}$. Examples of problems whose natural linear programming relaxation is $P$ (for some matrix $A$ and integer $b$) include the 2-edge-connected multigraph problem, Steiner tree problem, tree augmentation problem, and many others.

\begin{restatable}{theorem}{tightcuts}
	\label{tightcuts}
	Let $x\in P$, there exists $\theta\in [0,1]^k$, where $\sum_{i=1}^{k}\theta_i = 1$ and $\tilde{x}^i \in S$ for $i=1,\ldots,k$ such that \begin{itemize}
		\item for $\ell\in \{1,\ldots,n\}$, if $x_\ell =0$, then $\tilde{x}^i_\ell=0$ for $i =1,\ldots,k$, i.e. $\tilde{x}^i$ is in the support of $x$,
		\item we have $A_j (\sum_{i=1}^{k}\theta_i\tilde{x}^i)\leq C\cdot A_j x$, for $j$ such that $A_j x =b$,  
	\end{itemize}
	if and only if $C\geq g$.
\end{restatable}

This means in order to prove an upper bound on the integrality gap of a  bounded covering problem, we need to show there is a convex combination of integer feasible points that is ``cheap'' on all tight cuts. Notice that Theorem \ref{CV} requires the certificate convex combination to be ``cheap" on every single variable.

@@

Known polyhedral structure makes it easier to study integrality gaps for such problems. Carr and Ravi \cite{carr-ravi} introduced fundamental extreme points. A point $x$ in Held-Karp relaxation for TSP (or 2EC; they have the same relaxation) is a point whose support of $x$, namely $G_x$ satisfies the following: i)  $G_x$ is a cubic graph, ii) in $G_x$ there is exactly one edge with $x_e=1$ incident to each node iii) The fractional edges of $G_x$ form a Hamiltonian cycle.  We say a fundamental extreme point (FEP) is {\em order $k$} if there are $k$ nodes on this Hamiltonian cycle. An FEP of order $k$ could represent an instance with many more than $k$ vertices. Carr and co-authors~\cite{CV,carr-ravi,BC11} proved that showing that $Cx$ is a convex combination of tours (resp. 2-edge-connected multigraphs) for all fundamental extreme points is equivalent to proving that the integrality gap for TSP (resp. 2EC) is bounded above by $C$. We use fundamental extreme point to create the ``hardest'' LP solutions to decompose.

@@

\paragraph{Contributions}  The paper has the following contributions: 
\begin{itemize}
\item We give a simple algorithm, DomToIP, that can prove a binary formulation’s integrality gap is unbounded or if not provide a feasible integer solution.  Someone formulating a first MILP for a new problem can test it with DomToIP.  If the algorithm ever fails in finding a feasible solution, the MILP has an unbounded gap.
\item We give an algorithm, Fractional Decomposition Tree (FDT), to construct the convex decomposition in the Carr-Vempala theorem, perhaps scaling by a factor larger than the integrality gap.  Each step of this algorithm provides a {\em lower bound} on the instances integrality gap. This also provides a lower bound on the approximation factor of any LP-based approximation algorithm using this formulation. The overall approximation factor of the FDT algorithm is an upper bound on the integrality gap for that specific instance.
\item For a special set of problems related to TSP, where there is a notion of a fundamental extreme point and long-running attempts to exactly determine the integrality gap of classic formulations, experimental analysis with FDT can help give some intuition about which bound(s) is/are likely to be loose.  Computing on fundamental extreme points is a way to experimentally characterize the gap upper bound.  There is no guarantee. Still, this can help direct theoretical analysis in the most promising direction.  For instance for 2EC, FDT gives good approximate solutions, better than the best current competitor (Christofides' algorithm).

\item A stronger theoretical result on integrality gap that can be potentially employed to obtain improved upper bounds on the integrality gap of bounded covering problems.  
\end{itemize}

\iffalse{
Finally we give a stronger characterization of integrality gap than that in Theorem \ref{CV} for bounded covering problems. For this purpose assume $P= \{x\in \mathbb{R}^n_{\geq 0}: Ax\geq b\textbf{1}, x \leq b\textbf{1}\}$, $S= P\cap \mathbb{Z}^n$, and $g= \max_{c\geq 0} \frac{\min_{x\in S}cx}{\min_{x\in P}cx}$. Examples of such problems include the 2-edge-connected multigraph problem, the tree augmentation problem, and many others.

@@

This means in order to prove an upper bound on the integrality gap of a covering problem, we need to show there is a convex combination of integer feasible points that is ``cheap'' on all tight cuts. Notice that Theorem \ref{CV} requires the certificate convex combination to be ``cheap" on every single variable.
\subsection{Notations}
For vectors $x,y\in \mathbb{R}_{n}$ we say $x$ dominates $y$ if $x_i\geq y_i$ for $i= 1,\ldots,n$. For $m\times n$ matrix $A$, let $A_j$ be the $j$-th row of $A$ and $A^j$ be the $j$-th column of $A$. Let $\textbf{1}$ be the vector of all ones of a suitable dimension. For a set $S$ of vectors in $\mathbb{R}_{n}$, $\conv(S)$ is the convex hull of all the points in $S$.
}\fi
